---
title: "Predicting Glucose Levels in Diabetes Patients"
author: "STATS 101A Group 35"
date: "2023-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readr")
library("tidyr")
library("dplyr")
library("car")
library("leaps")
library("vtable")
library("ggplot2")
library("gridExtra")
```

## Importing Dataset

```{r}
diabetes <- read.csv("diabetes.csv")
```

```{r}
head(diabetes)
```

```{r}
count_zeros <- function(x) {
  sum(x == 0, na.rm = TRUE)
}

zero_counts <- sapply(diabetes[, 0:8], count_zeros)
data.frame(column_names = names(zero_counts), zero_count = zero_counts)
```

```{r}
diabetes <- diabetes %>% 
  filter(Glucose != 0, BloodPressure != 0, Insulin != 0, BMI != 0, DiabetesPedigreeFunction != 0, Age != 0, Pregnancies != 0)

attach(diabetes)
```

```{r}
zero_counts <- sapply(diabetes[, 0:8], count_zeros)
data.frame(column_names = names(zero_counts), zero_count = zero_counts)
```

```{r}
plot_list <- lapply(names(diabetes), function(column) {
  ggplot(diabetes, aes_string(x = column)) +
    geom_density() +
    labs(title = column,
         x = column,
         y = "Density") +
    theme_minimal()
})

combined_plots <- do.call(grid.arrange, c(plot_list, ncol = 3))
print(combined_plots)
```

## Initial Model Analysis

```{r}
model <- lm(Glucose ~ Pregnancies+Insulin+BMI+SkinThickness+BloodPressure+DiabetesPedigreeFunction+Age+Outcome)
```

#### Scatterplot and Correlation Matrices

```{r}
pairs(diabetes, cex = 0.1)
cor(diabetes)
```

#### Summary of Each Variable

```{r}
st(diabetes)
```


#### Model Summary

```{r}
summary(model)
```

#### ANOVA

```{r}
anova(model)
```

#### Diagnostic Plots

```{r}
par(mfrow = c(2,2))
plot(model)
```

#### Leverages and Outliers

```{r}
hats <- hatvalues(model)
leverages <- hats[hats > 2 * mean(hats)]
outliers <- hats[abs(rstandard(model)) >= 2]

outliers[names(outliers) %in% names(leverages)]
```

#### Multicolinearity Validation

```{r}
vif(model)
```

#### Standardized Residual Plots

```{r}
stanRes <- rstandard(model)
par(mfrow = c(2, 3))
plot(Insulin, stanRes, ylab = "Standardized Residuals")
plot(BMI, stanRes, ylab = "Standardized Residuals")
plot(BloodPressure, stanRes, ylab = "Standardized Residuals")
plot(DiabetesPedigreeFunction, stanRes, ylab = "Standardized Residuals")
plot(Age, stanRes, ylab = "Standardized Residuals")
```

#### Added Variable Plots

```{r}
par(mfrow=c(2, 3))
avPlot(model, variable = "BloodPressure", ask = F)
avPlot(model, variable = "Insulin", ask = F)
avPlot(model, variable = "BMI", ask = F)
avPlot(model, variable = "DiabetesPedigreeFunction", ask = F)
avPlot(model, variable = "Age", ask = F)
```

## Model Transformation

#### Box-Cox Transformation

```{r}
summary(powerTransform(cbind(Insulin, BMI, BloodPressure, DiabetesPedigreeFunction, Age) ~ 1))
```

#### Inverse-Response Plot

```{r}
tInsulin <- log(Insulin)
tBMI <- log(BMI)
tBloodPressure <- BloodPressure
tDiabetesPedigreeFunction <- log(DiabetesPedigreeFunction)
tAge <- Age^(-2)

t_model <- lm(Glucose ~ tInsulin+tBMI+tBloodPressure+tDiabetesPedigreeFunction+tAge)
invResPlot(t_model)
```

## Transformed Model Analysis

```{r}
tGlucose <- log(Glucose)
t_model <- lm(tGlucose ~ tInsulin+tBMI+tBloodPressure+tDiabetesPedigreeFunction+tAge)
```

#### Model Summary

```{r}
summary(t_model)
```

#### ANOVA

```{r}
anova(t_model)
```

#### Diagnostic Plots

```{r}
par(mfrow = c(2, 2))
plot(t_model)
```

#### Leverages and Outliers

```{r}
t_hats <- hatvalues(t_model)
t_leverages <- t_hats[hats > 2 * mean(t_hats)]
t_outliers <- t_hats[abs(rstandard(t_model)) >= 2]

t_outliers[names(t_outliers) %in% names(t_leverages)]
```

#### Multicolinearity Validation

```{r}
vif(t_model)
```

#### Standardized Residual Plots

```{r}
t_stanRes <- rstandard(t_model)
par(mfrow = c(2, 3))
plot(Insulin,  t_stanRes, ylab = "Standardized Residuals")
plot(BMI,  t_stanRes, ylab = "Standardized Residuals")
plot(BloodPressure,  t_stanRes, ylab = "Standardized Residuals")
plot(DiabetesPedigreeFunction,  t_stanRes, ylab = "Standardized Residuals")
plot(Age,  t_stanRes, ylab = "Standardized Residuals")
```

## Variable Selection

#### All Possible Subsets Method

```{r}
summary(regsubsets(as.matrix(cbind(tInsulin, tBMI, tBloodPressure, tDiabetesPedigreeFunction, tAge)), tGlucose))
```

```{r}
om1 <- lm(tGlucose ~ tInsulin)
om2 <- lm(tGlucose ~ tInsulin + tAge)
om3 <- lm(tGlucose ~ tInsulin + tBloodPressure + tAge)
om4 <- lm(tGlucose ~ tInsulin + tBloodPressure + tDiabetesPedigreeFunction  + tAge)
om5 <- lm(tGlucose ~ tInsulin + tBMI + tBloodPressure + tDiabetesPedigreeFunction  + tAge)
possible_models <- list(om1, om2, om3, om4, om5)
p <- 1:5
n <- nrow(diabetes)

criterias_breakdown <- matrix(0, nrow = 5, ncol = 5)
colnames(criterias_breakdown) <- c("Size", "Adj R2", "AIC", "AICc", "BIC")

# Adjusted R-squared
adj_R2 <- sapply(possible_models, function(mod) { summary(mod)$adj.r.squared })

# AIC
AIC <- sapply(possible_models, function(mod) { extractAIC(mod)[2] })

# AICc
AICc <- rep(0, 5)
for (i in p) {
  AICc[i] <- AIC[i] + ((2 * p[i]) * (p[i] + 2) * (p[i] + 3)) / (n - p[i] - 1)
}

# BIC
BIC <- sapply(possible_models, function(mod) { extractAIC(mod, k = log(n))[2] })

criterias_breakdown[, 1:5] <- c(p, adj_R2, AIC, AICc, BIC)
criterias_breakdown
```

```{r}
summary(om2)
```

```{r}
summary(om3)
```

## Final Model

```{r}
final_mod <- lm(tGlucose ~ tInsulin + tBloodPressure + tAge)
summary(final_mod)
```

```{r}
detach("diabetes")
```




